Tunable Virtual IMU Frame by Averaging Multiple Non-collocated IMUs

### 2024-09-12 paper/main.tex
Addressed small fixes from Tim.

% \title{Weighted Averaging of Multiple Non-collocated Inertial Measurement Units}
\title{Tunable VIMU frame from averaging multiple Non-collocated IMUs}

\author{\IEEEauthorblockN{1\textsuperscript{st} Yizhou Gao}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Timothy D. Barfoot}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
}

@@

\begin{abstract}
Multi-IMU fusion has been an effective way of improving MEMS IMU accuracy. In this article, we discuss an averaging technique for multiple IMUs with physical separation. We also present methods of selecting weighting terms such that we can minimize the uncertainty in fused sensor output and select the placement of the reference frame of the combined virtual IMU (VIMU). We tested our method in simulation and validated it on a real dataset. The results show that the averaging technique works for IMUs with large separation and performance gain is observed in simulation compared to using only a single IMU.
\end{abstract}

\begin{IEEEkeywords}
IMU, multi-IMU, sensor fusion, VINS
\end{IEEEkeywords}

@@

An inertial measurement unit combining a gyroscope and accelerometer that provides measurement of angular velocity and linear acceleration has wide application in robotics and UAS. By integrating the measurement, one can track the relative orientation and position of the system to certain degree of accuracy within reasonable time frame \cite{tang2022_preintegration}. However due to the interoceptive nature of an IMU, it is often paired with additional sensors such as camera, lidar, or GNSS to reduce drift \cite{alaba2024_gps_and_imu, shan2020_liosam, he2020_visual_imu_and_gps}. In general, there are two main approaches to the tracking problem with an IMU: filter-based and graph optimization with pre-integrated IMU measurements.

In the filter-based approach, the IMU measurement is used as process input to an Extended Kalman Filter in the propagation step. Then the measurements from other exteroceptive sensors are used for state correction in the update step. One such example is the MSCKF (Multi-State Constraint Kalman Filter) proposed in \cite{Anastasios2007_MSCKF}, which uses the IMU for propagation and a camera for correction. It proposes a measurement model that directly relates two robot poses based on the landmark observation in the camera but independent from the landmark position itself. This allows for efficient tracking without including the positions of landmarks as part of the state vector as in traditional EKF-SLAM.

For the graph-optimization approach, such as VINS-MONO \cite{qin2018_vins-mono}, a pose graph is constructed by solving tightly coupled, nonlinear optimization problem with preintegrated IMU measurement and camera observations. Pose-graph optimization is run at the end to ensure global consistency.

Among many sensor configurations with an IMU, monocular camera plus IMU is one of the minimal configuration that is well-studied \cite{qin2018_vins-mono, 10616216}. Also IMU has been used heavily as a complementary sensor for GPS in GNSS-denied environment \cite{1008998, 8987949}. Based on these minimal configurations, various extension has been made. UMS-VINS \cite{jiang2023_UMS-VINS} provides a unified framework for using monocular and stereo cameras in visual-inertial odometry. Also the MSCKF framework \cite{Anastasios2007_MSCKF} works for multiple cameras with an IMU. There are also works that aim to provide a complete framework to incorporate multiple different sensors in one unified system \cite{10587194}.

@@

Despite works mentioned above, most systems and applications still only use single IMU in the estimation pipeline. The vehicle frame is choose to be at the frame of reference of the IMU to avoid introducing additional kinematics terms from the lever arm effect.

In this article we present that through weighted averaging of multiple non-collocated IMUs, we can choose arbitrary VIMU frame of reference within solvable domain.

@@

One category of approach is to perform fusion of multiple IMUs before using the measurement in inertial navigation system (INS). For example, \cite{jafari2014_PEM} uses Prediction Error Minimization method to model noise of IMUs and use Kalman filter to estimate the optimal process signal. \cite{xue2023} proposed to use Kalman filter to estimate the angular velocity measured by multiple non-orthogonal gyroscopes. Instead of using Kalman filter, an easier approach would be to directly average multiple gyroscopes after aligned under a common virtual IMU frame as discussed in \cite{waegli2008, patel2022_multi-imu, Colomina2004REDUNDANTIF}.

However, averaging of multiple accelerometers was not considered due to the additional acceleration terms from rotation (lever arm effect), unless they are closely placed and the separations in between are ignored.

@@

Typically state estimation problem with a IMU involves tracking the orientation, linear velocity and linear acceleration of the vehicle (commonly defined at the IMU). With additional IMUs, additional state variables like linear acceleration, angular velocity and angular acceleration are added. This allows using the full mechanization model of accelerometer as measurement update in Kalman filter \cite{Bancroft2011DataFA, Beaudoin2018_satelite}.

@@

Another approach is to track the state of each IMU individually and associate them with geometric constraints \cite{waegli2008, Beaudoin2018_satelite}. Using multiple IMUs and cameras, \cite{Eckenhoff2021_MIMC-VINS} proposed to track all stacked states of base and auxillary IMUs using MSCKF \cite{Anastasios2007_MSCKF}. Assuming the extrinsic calibrations of all IMUs are known and fixed, relative geometric constraints is imposed between states of different IMUs in the form of measurement update in addition to camera observations.

@@

Last common approach would be to have individual state estimator for each IMU system each performing local propagation and correction. Simple state fusion is performed with each estimator's estimate \cite{Bancroft2011DataFA, patel2022_multi-imu}.

@@

As presented, multiple approaches have being proposed. \ref{constraint} would be computationally expensive if number of IMU are large as it has many states all in one Kalman filter. The averaging approach in \ref{VIMU} would be the simplest to implement and doesn't significantly alter the existing single IMU framework. However, due to the lever arm effect, averaging multiple accelerometers that are physically separated is not visited. In \ref{methodology} we present that it is still possible to average the accelerometers despite the separation and give a tunable VIMU frame as a result.

@@

In this discussion, we use the following four frames of reference. 1) Inertial frame $\bm{\mathcal{F}}_i$. 2) Body frame $\bm{\mathcal{F}}_b$ is for expressing sensor extrinsic calibration relative to some mechanical datum. 3) Sensor frame $\bm{\mathcal{F}}_s$ is the local frame of reference at each sensor. 4) Vehicle frame $\bm{\mathcal{F}}_v$ is for tracking vehicle pose in state estimation pipeline. In later section we use vehicle frame interchangeably with VIMU frame.

For any point $\textbf{p}$ or vector we define as $\textbf{r}_i^{pi}$ in inertial frame $\mathcal{F}_i$. For example, the vehicle position in inertial frame thus become $\textbf{r}_i^{vi}$.

When referring to the pose of a vehicle, it is the tuple of $\{\textbf{C}_{iv}, \textbf{r}_i^{vi}\}$ which forms the transformation matrix as follow,
$$
\textbf{T}_{iv} = \left[\begin{matrix}
    \textbf{C}_{iv} & \textbf{r}_i^{vi} \\
    \textbf{0} & 1
\end{matrix}\right]
$$

The pose can be used to transform point $\textbf{r}_v^{pv}$ expressed in vehicle frame $\bm{\mathcal{F}}_v$ into inertial frame $\bm{\mathcal{F}}_i$ as $\textbf{r}_i^{pi}$:
$$
\left[\begin{matrix} \textbf{r}_i^{pi} \\ 1 \end{matrix}\right]
    = \textbf{T}_{iv} \left[\begin{matrix} \textbf{r}_v^{pv} \\ 1 \end{matrix}\right]
\quad \text{or} \quad
\textbf{r}_i^{pi} = \textbf{C}_{iv} \textbf{r}_v^{pv} + \textbf{r}_i^{vi}
$$

In this document $*^\wedge$ is specifically used for expressing a $\mathbb{R}^3$ as skew symmetric matrix.
$$
\bm{\omega}^\wedge = \left[\begin{matrix}
    0 & -\omega_2 & \omega_1 \\
    \omega_2 & 0 & -\omega_3 \\
    -\omega_1 & \omega_3 & 0
\end{matrix}\right]
$$

If a quantity is in the form $x(t)$ then it is a time-varying quantity, otherwise it is time-invariant.

@@

The state propagation using IMU measurement follows the rigid-body kinematics in world fixed frame.
\begin{equation}
\begin{align}
    \dot{\textbf{C}}(t) &= \textbf{C}(t) \bm{\omega}(t)^\wedge \\
    \dot{\textbf{v}}(t) &= \textbf{C}(t) \textbf{a}(t) - \textbf{g} \\
    \dot{\textbf{p}}(t) &= \textbf{v}(t)
\end{align}
\end{equation}

Here $\textbf{C}$ is short for $\textbf{C}_{iv}$, $\textbf{v}$ is short for $\textbf{v}_i^{vi}$, and $\textbf{p}$ is for the position of the vehicle $\textbf{r}_i^{vi}$. These three formed the 9-dof state variables.

In addition, $\textbf{g}$ is the acceleration from gravity and $\bm{\omega}$ and $\textbf{a}$ are the true angular velocity and true linear acceleration in vehicle frame.

@@

For the angular velocity measurement from gyroscope we have the following measurement model.

\begin{equation}
    \textbf{y}_\omega(t) = \bm{\omega}(t) + \textbf{b}_\omega(t) + \textbf{n}_\omega(t)
\end{equation}

Where $\textbf{b}_\omega$ is the bias of gyroscope measurement due to random walk and $\textbf{n}_\omega$ is the additive white noise of the measurement. Similarly we have the following measurement model for accelerometer at vehicle frame. Here $\textbf{b}_a$ is the bias of accelerometer and $\textbf{n}_a$ is the noise of the acceleration measurement.

\begin{equation}
    \textbf{y}_a(t) = \textbf{a}(t) + \textbf{b}_a(t) + \textbf{n}_a(t)
\end{equation}

Bias drifts of IMU are modeled as random walk with the bias drift rate sampled from a Gaussian noise.
\begin{equation}
\begin{split}
    \dot{\textbf{b}}_\omega(t) &= \textbf{w}_\omega(t) \\
    \dot{\textbf{b}}_a(t) &= \textbf{w}_a(t)
\end{split}
\end{equation}

Due to the presence of biases, it is estimated as part of the state vector and subtracted from the raw measurement to get the unbiased measurement for propagation.
\begin{equation}
\begin{split}
    \hat{\bm{\omega}}(t) = \textbf{y}_\omega(t) - \hat{\textbf{b}}_\omega(t) \\
    \hat{\textbf{a}}(t)  = \textbf{y}_a(t) - \hat{\textbf{b}}_a(t)
\end{split}
\end{equation}

For IMU not at vehicle frame, measurement are in each IMU's local frame and we have additional terms in acceleration measurement due to lever arm effect.
\begin{equation}
    \textbf{y}_{\omega,i}(t) = \textbf{C}_{i}^{-1} \bm{\omega}(t) + \textbf{b}_{\omega, i}(t) + \textbf{n}_{\omega,i}(t)
\end{equation}
\begin{equation}
\begin{split}
    \textbf{y}_{a,i}(t) &= \textbf{C}_i^{-1} \left[ \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \right] \\
    &+ \textbf{b}_{a,i}(t) + \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

Here we have $\textbf{C}_i$ and $\textbf{r}_i$ short for $\textbf{C}_{vs}$ and $\textbf{r}_v^{sv}$ defining the pose $\textbf{T}_{vs}$ of the $i$-th IMU in vehicle frame.

@@

First we align the gyroscope measurement with the vehicle frame.
\begin{equation}
    \textbf{C}_{i} \textbf{y}_{\omega,i}(t) = \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega, i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t)
\end{equation}

The weighted average ($\sum_i{w_i} = 1$) of all gyroscope is simply,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_\omega(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_\omega(t)} \\
    &= \sum_i{w_i \left( \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega,i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t) \right)} \\
    &= \bm{\omega}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{\omega,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{\omega,i}(t)}
\end{split}
\end{equation}

We see that the averaged measurement can be summarized as,
\begin{equation}
    \bar{\textbf{y}}_\omega(t) = \bm{\omega}(t) + \bar{\textbf{b}}_\omega(t) + \bar{\textbf{n}}_\omega(t)
\end{equation}

@@

Align all accelerometer measurements to vehicle frame we have,
\begin{equation}\label{eqn_accel}
\begin{split}
    \textbf{C}_{i} \textbf{y}_{a,i}(t) &= \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \\
    &+ \textbf{C}_{i} \textbf{b}_{a,i}(t) + \textbf{C}_{i} \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

Summing equation (\ref{eqn_accel}) of all accelerometers gives,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_a(t)} \\
    &= \textbf{a}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \sum_i{w_i \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i)} + \sum_i{w_i \bm{\alpha}(t)^\wedge \textbf{r}_i} \\
\end{split}
\end{equation}

Using the associate rule $\textbf{a}^\wedge \textbf{b} + \textbf{a}^\wedge \textbf{c} = \textbf{a}^\wedge \left(\textbf{b} + \textbf{c}\right)$, we have,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_a(t)} \\
    &= \textbf{a}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \bm{\omega}(t)^\wedge \left(\bm{\omega}(t)^\wedge \left( \sum_i{w_i \textbf{r}_i} \right) \right) + \bm{\alpha}(t)^\wedge \left( \sum_i{w_i\textbf{r}_i} \right)\\
\end{split}
\end{equation}

We see that if $\sum_i{w_i \textbf{r}_i} = \textbf{0}$ then we can eliminate the last two terms from lever arm effect. This gives,
\begin{equation}
    \bar{\textbf{y}}_a(t) = \textbf{a}(t) + \bar{\textbf{b}}_a(t) + \bar{\textbf{n}}_a(t)
\end{equation}

@@

As we can see, if we choose the vehicle frame and weights of accelerometers such that $\sum_i{w_i \textbf{r}_i} = \textbf{0}$, then we can ignore the separation of all IMUs and fuse the accelerometer measurement through averaging.

Since gyroscope measurement is location independent, we can choose different weights for averaging gyroscope and accelerometers.

Since all biases are slow-varying, the combined biases are also slow-varying. So instead of tracking individual biases, only the combined biases are needed. When receiving new set of IMU measurement, we estimate the angular velocity and linear acceleration with the update equations below.
\begin{equation}
\begin{split}
    \hat{\bm{\omega}}(t) = \bar{\textbf{y}}_\omega(t) - \hat{\bar{\textbf{b}}}_\omega(t) \\
    \hat{\textbf{a}}(t)  = \bar{\textbf{y}}_a(t) - \hat{\bar{\textbf{b}}}_a(t)
\end{split}
\end{equation}

The combined measurement is expected to have slower drift and the estimate is expected to be less noisy as well \cite{patel2022_multi-imu}. If having $n$ identical IMUs and equal weight the standard deviation of the combined noise or drift rate is expected to be,
\begin{equation}
    \bar{\sigma} = \frac{\sigma}{\sqrt{n}}
\end{equation}

\subsection{Selecting Weight for Minimal Uncertainty}\label{solve_weight_by_noise}

When selecting weight for gyroscopes, or for accelerometers under the condition that the selection of vehicle frame is flexible, we can minimize the output noise or bias drift rate. Below we take the bias drift as an example but similar idea can be applied to the output noise.

The combined drift rate is as follow. Here we drop the subscript as it is the same for both gyroscope drift and accelerometer drift.
\begin{equation}
    \dot{\bar{\textbf{b}}}(t) = \sum_i{w_i \textbf{C}_i \textbf{w}_i(t)}
\end{equation}

The covariance of the combined drift rate would be the weighted sum of the noise covariance of each IMU.
\begin{equation}
    \bar{\bm{\Sigma}} = \sum_i{w_i^2 \bm{\Sigma}_i}
\end{equation}

One can then solve for the weights that minimize the trace or the sum of the singular value of the covariance.

>>>>

% \title{Weighted Averaging of Multiple Non-collocated Inertial Measurement Units}
\title{Tunable Virtual IMU Frame by Averaging Multiple Non-collocated IMUs}

\author{\IEEEauthorblockN{Yizhou Gao}
\IEEEauthorblockA{\textit{Institute for Aerospace Studies} \\
\textit{University of Toronto}\\
Toronto, Canada \\
yizhou.gao@robotics.utias.utoronto.ca}
\and
\IEEEauthorblockN{Timothy D. Barfoot}
\IEEEauthorblockA{\textit{Institute for Aerospace Studies} \\
\textit{University of Toronto}\\
Toronto, Canada \\
tim.barfoot@utoronto.ca}}

@@

\begin{abstract}
Multi-IMU fusion has been an effective way of improving MEMS IMU accuracy. In this article, we discuss an averaging technique for multiple IMUs with physical separation. We also present methods of selecting weighting terms such that we can minimize the uncertainty in fused sensor output and select the placement of the reference frame of the combined virtual IMU (VIMU). The VIMU can be placed to be concident with, for example, a camera frame or GNSS frame thereby offering a quality-of-life improvement for users. We tested our method in simulation and validated it on a real dataset. The results show that the averaging technique works for IMUs with large separation and performance gain is observed in both simulation and real experiment compared to using only a single IMU.
\end{abstract}

\begin{IEEEkeywords}
IMU, multi-IMU, VIMU, sensor fusion, VINS
\end{IEEEkeywords}

@@

An inertial measurement unit combining a gyroscope and accelerometer that provides measurement of angular velocity and linear acceleration has wide application in robotics. By integrating the measurement, one can track the relative orientation and position of the system to certain degree of accuracy within a reasonable time frame \cite{tang2022_preintegration}. However, due to the interoceptive nature of an IMU, it is often paired with additional sensors such as camera, lidar, or GNSS to reduce drift \cite{alaba2024_gps_and_imu, shan2020_liosam, he2020_visual_imu_and_gps}.

% In general, there are two main approaches to the tracking problem with an IMU: filter-based and graph optimization with pre-integrated IMU measurements.

% In the filter-based approach, the IMU measurement is used as process input to an Extended Kalman Filter in the propagation step. Then the measurements from other exteroceptive sensors are used for state correction in the update step. One such example is the MSCKF (Multi-State Constraint Kalman Filter) proposed in \cite{Anastasios2007_MSCKF}, which uses the IMU for propagation and a camera for correction. It proposes a measurement model that directly relates two robot poses based on the landmark observation in the camera but independent from the landmark position itself. This allows for efficient tracking without including the positions of landmarks as part of the state vector as in traditional EKF-SLAM.

% For the graph-optimization approach, such as VINS-MONO \cite{qin2018_vins-mono}, a pose graph is constructed by solving tightly coupled, nonlinear optimization problem with preintegrated IMU measurement and camera observations. Pose-graph optimization is run at the end to ensure global consistency.

Among many sensor configurations with an IMU, monocular camera plus IMU is one of the minimal configurations that is well-studied \cite{qin2018_vins-mono, 10616216}. Also IMU has been used heavily as a complementary sensor for GPS in GNSS-denied environment \cite{1008998, 8987949}. Based on these minimal configurations, various extensions have been made. UMS-VINS \cite{jiang2023_UMS-VINS} provides a unified framework for using monocular and stereo cameras in visual-inertial odometry. Also the MSCKF framework \cite{Anastasios2007_MSCKF} works for multiple cameras with an IMU. There are also works that aim to provide a complete framework to incorporate multiple different sensors in one unified system \cite{10587194}.

@@

Despite the works mentioned above, most systems and applications still only use a single IMU in the estimation pipeline. The vehicle frame is chosen to be at the frame of reference of the IMU to avoid introducing additional kinematics terms from the lever-arm effect. If such IMU frame differ from a frame of reference of concern, additional transformation is needed after the estimation pipeline.

In this article, we show that through a weighted averaging of multiple non-collocated IMUs, we can choose an arbitrary VIMU frame of reference that gives quality-of-life improvement and offering additional benefit of reduced noise and bias drift.

@@

One approach is to perform fusion of multiple IMUs before using the measurement in inertial navigation system (INS). For example, \cite{jafari2014_PEM} uses Prediction Error Minimization method to model noise of IMUs and uses a Kalman filter to estimate the optimal process signal. \cite{xue2023} proposed to use a Kalman filter to estimate the angular velocity measured by multiple non-orthogonal gyroscopes. Instead of using a Kalman filter, an easier approach would be to directly average multiple gyroscopes after aligned under a common virtual IMU frame as discussed in \cite{waegli2008, patel2022_multi-imu, Colomina2004REDUNDANTIF}.

However, averaging of multiple accelerometers was not considered due to the additional acceleration terms from rotation (lever arm effect), unless they are closely placed and the separations in between are ignored. In section \ref{methodology} we will show how it is possible to average multiple accelerometers even with separations.

@@

Typically, a state estimation problem with a IMU involves tracking the orientation, linear velocity, and linear acceleration of the vehicle (commonly defined at the IMU). With additional IMUs, additional state variables such as linear acceleration, angular velocity, and angular acceleration are added. This allows using the full mechanization model of accelerometer as the measurement update in the Kalman filter \cite{Bancroft2011DataFA, Beaudoin2018_satelite}. On the contrary, our approach only involves averaging the measurements thus keeping the computational overhead minimal.

@@

Another approach is to track the state of each IMU individually and associate them with geometric constraints \cite{waegli2008, Beaudoin2018_satelite}. Using multiple IMUs and cameras, \cite{Eckenhoff2021_MIMC-VINS} proposed to track all stacked states of base and auxillary IMUs using MSCKF \cite{Anastasios2007_MSCKF}. Assuming the extrinsic calibrations of all IMUs are known and fixed, relative geometric constraints are imposed between states of different IMUs in the form of measurement updates in addition to camera observations. Our approach allows fusing the information from multiple IMUs without the need of introducing the extra geometric constraints in the state estimation pipeline.

@@

The last common approach is to have an individual state estimator for each IMU system each performing local propagation and correction. Simple state fusion is performed with each estimator's estimate \cite{Bancroft2011DataFA, patel2022_multi-imu}. Our approach avoids the need to run several state estimators which can be expensive.

@@

As presented, multiple approaches have being proposed. \ref{constraint} would be computationally expensive if the number of IMUs is large as it has many states all in one Kalman filter. The averaging approach in \ref{VIMU} would be the simplest to implement and does not significantly alter the existing single IMU framework. However, due to the lever-arm effect, averaging multiple accelerometers that are physically separated is not visited. We will show that it is still possible to average the accelerometers despite the separation and give a tunable VIMU frame as a result.

@@

In this discussion, we use the following four frames of references: 1) inertial frame $\bm{\mathcal{F}}_i$, 2) body frame $\bm{\mathcal{F}}_b$ is for expressing sensor extrinsic calibration relative to some mechanical datum, 3) sensor frame $\bm{\mathcal{F}}_s$ is the local frame of reference at each sensor, 4) VIMU frame or Vehicle frame $\bm{\mathcal{F}}_v$ is for tracking vehicle pose in state estimation pipeline.

For any point $\textbf{p}$ or vector we define as $\textbf{r}_i^{pi}$ in the inertial frame $\bm{\mathcal{F}}_i$. For example, the vehicle position in the inertial frame thus becomes $\textbf{r}_i^{vi}$.

When referring to the pose of a vehicle, it is the tuple of $\{\textbf{C}_{iv}, \textbf{r}_i^{vi}\}$, which forms the transformation matrix as follows:
$$
\textbf{T}_{iv} = \left[\begin{matrix}
    \textbf{C}_{iv} & \textbf{r}_i^{vi} \\
    \textbf{0} & 1
\end{matrix}\right].
$$

The pose can be used to transform point $\textbf{r}_v^{pv}$ expressed in vehicle frame $\bm{\mathcal{F}}_v$ into inertial frame $\bm{\mathcal{F}}_i$ as $\textbf{r}_i^{pi}$:
$$
\left[\begin{matrix} \textbf{r}_i^{pi} \\ 1 \end{matrix}\right]
    = \textbf{T}_{iv} \left[\begin{matrix} \textbf{r}_v^{pv} \\ 1 \end{matrix}\right]
\quad \text{or} \quad
\textbf{r}_i^{pi} = \textbf{C}_{iv} \textbf{r}_v^{pv} + \textbf{r}_i^{vi}.
$$

In this document $^\wedge$ is specifically used for rendering an element of $\mathbb{R}^3$ as a skew-symmetric matrix:
$$
\bm{\omega}^\wedge = \left[\begin{matrix}
    0 & -\omega_2 & \omega_1 \\
    \omega_2 & 0 & -\omega_3 \\
    -\omega_1 & \omega_3 & 0
\end{matrix}\right].
$$

If a quantity is in the form $x(t)$, then it is a time-varying quantity, otherwise it is time-invariant.

@@

The state propagation using IMU measurements follows the rigid-body kinematics in world fixed frame:
\begin{equation}
\begin{align}
    \dot{\textbf{C}}(t) &= \textbf{C}(t) \bm{\omega}(t)^\wedge, \\
    \dot{\textbf{v}}(t) &= \textbf{C}(t) \textbf{a}(t) - \textbf{g}, \\
    \dot{\textbf{p}}(t) &= \textbf{v}(t).
\end{align}
\end{equation}

Here $\textbf{C}$ is short for $\textbf{C}_{iv}$, $\textbf{v}$ is short for $\textbf{v}_i^{vi}$, and $\textbf{p}$ is for the position of the vehicle $\textbf{r}_i^{vi}$. These three form a 9-DOF state.

In addition, $\textbf{g}$ is the acceleration from gravity and $\bm{\omega}$ and $\textbf{a}$ are the true angular velocity and true linear acceleration in the vehicle frame.

@@

For the angular velocity measurement from a gyroscope we have the following measurement model:

\begin{equation}
    \textbf{y}_\omega(t) = \bm{\omega}(t) + \textbf{b}_\omega(t) + \textbf{n}_\omega(t),
\end{equation}

\noindent where $\textbf{b}_\omega$ is the bias of gyroscope measurement due to random walk and $\textbf{n}_\omega$ is the additive white noise of the measurement. Similarly, we have the following measurement model for an accelerometer at the vehicle frame:

\begin{equation}
    \textbf{y}_a(t) = \textbf{a}(t) + \textbf{b}_a(t) + \textbf{n}_a(t),
\end{equation}

\noindent here $\textbf{b}_a$ is the bias of accelerometer and $\textbf{n}_a$ is the noise of the acceleration measurement.

Bias drifts of IMU are modeled as random walk with the bias drift rate sampled from a Gaussian noise:
\begin{equation}
\begin{split}
    \dot{\textbf{b}}_\omega(t) &= \textbf{w}_\omega(t), \\
    \dot{\textbf{b}}_a(t) &= \textbf{w}_a(t).
\end{split}
\end{equation}

\noindent Due to the presence of biases, it is estimated as part of the state vector and subtracted from the raw measurement to get the "unbiased" measurement for propagation:
\begin{equation}
\begin{split}
    \hat{\bm{\omega}}(t) = \textbf{y}_\omega(t) - \hat{\textbf{b}}_\omega(t), \\
    \hat{\textbf{a}}(t)  = \textbf{y}_a(t) - \hat{\textbf{b}}_a(t).
\end{split}
\end{equation}

For an IMU not located at the vehicle frame, measurements are in each IMU's local frame and we have additional terms in the acceleration measurement due to lever arm effect:
\begin{equation}
\begin{split}
    \textbf{y}_{\omega,i}(t) &= \textbf{C}_{i}^{-1} \bm{\omega}(t) + \textbf{b}_{\omega, i}(t) + \textbf{n}_{\omega,i}(t), \\
    \textbf{y}_{a,i}(t) &= \textbf{C}_i^{-1} \left[ \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \right] \\
    &+ \textbf{b}_{a,i}(t) + \textbf{n}_{a,i}(t).
\end{split}
\end{equation}

\noindent Here we have $\textbf{C}_i$ and $\textbf{r}_i$ short for $\textbf{C}_{vs}$ and $\textbf{r}_v^{sv}$ defining the pose $\textbf{T}_{vs}$ of the $i$-th IMU in the vehicle frame.

@@

First, we align the gyroscope measurements with the vehicle frame:
\begin{equation}
    \textbf{C}_{i} \textbf{y}_{\omega,i}(t) = \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega, i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t).
\end{equation}

\noindent The weighted average of all gyroscopes is simply,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_\omega(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_{\omega,i}(t)} \\
    &= \sum_i{w_i \left( \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega,i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t) \right)} \\
    &= \bm{\omega}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{\omega,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{\omega,i}(t)}
\end{split}
\end{equation}

\noindent Where $w_i$ are the weights with $\sum_i{w_i} = 1$.

We see that the averaged measurement can be summarized as
\begin{equation}
    \bar{\textbf{y}}_\omega(t) = \bm{\omega}(t) + \bar{\textbf{b}}_\omega(t) + \bar{\textbf{n}}_\omega(t),
\end{equation}

\noindent where $\bar{\textbf{b}}_\omega(t) = \sum_i{w_i \textbf{C}_{i} \textbf{b}_{\omega,i}(t)}$ and $\bar{\textbf{n}}_\omega(t) = \sum_i{w_i \textbf{C}_{i} \textbf{n}_{\omega,i}(t)}$ are the combined biases and noises.

@@

Aligning all accelerometer measurements to the vehicle frame, we have
\begin{equation}\label{eqn_accel}
\begin{split}
    \textbf{C}_{i} \textbf{y}_{a,i}(t) &= \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \\
    &+ \textbf{C}_{i} \textbf{b}_{a,i}(t) + \textbf{C}_{i} \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

\noindent Summing (\ref{eqn_accel}) over all accelerometers gives
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_{a,i}(t)} \\
    &= \textbf{a}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \sum_i{w_i \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i)} + \sum_i{w_i \bm{\alpha}(t)^\wedge \textbf{r}_i} \\
\end{split}
\end{equation}

\noindent Using the fact that $^\wedge$ is linear, we have
\begin{equation}\label{accel_avg_final}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_{a,i}(t)} \\
    &= \textbf{a}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \bm{\omega}(t)^\wedge \left(\bm{\omega}(t)^\wedge \left( \sum_i{w_i \textbf{r}_i} \right) \right) + \bm{\alpha}(t)^\wedge \left( \sum_i{w_i\textbf{r}_i} \right).
\end{split}
\end{equation}

\noindent We see that if $\sum_i{w_i \textbf{r}_i} = \textbf{0}$, then we can {\em eliminate} the last two terms from the lever-arm effect. This simplifies (\ref{accel_avg_final}) to
\begin{equation}
    \bar{\textbf{y}}_a(t) = \textbf{a}(t) + \bar{\textbf{b}}_a(t) + \bar{\textbf{n}}_a(t).
\end{equation}

@@

As we can see, if we choose the vehicle frame and weights of theaccelerometers such that $\sum_i{w_i \textbf{r}_i} = \textbf{0}$, then we can ignore the separation of all IMUs and fuse the accelerometer measurements through averaging. The result is an virtual IMU at the weighted centroid without any lever-arm effects.

Since gyroscope measurement is location independent, we can choose different weights for averaging gyroscopes and accelerometers.

Because all biases are slowly varying, the combined biases are also slowly varying. Therefore, instead of tracking individual biases, only the combined biases are needed. When receiving a new set of IMU measurements, we estimate the angular velocity and linear acceleration with the simple update equations below:
\begin{equation}
\begin{split}
    \hat{\bm{\omega}}(t) = \bar{\textbf{y}}_\omega(t) - \hat{\bar{\textbf{b}}}_\omega(t), \\
    \hat{\textbf{a}}(t)  = \bar{\textbf{y}}_a(t) - \hat{\bar{\textbf{b}}}_a(t).
\end{split}
\end{equation}

The combined measurements are expected to have slower drift and the estimate is expected to be less noisy as well \cite{patel2022_multi-imu}. If having $n$ identical IMUs and equal weights the standard deviation of the combined noise or drift rate is expected to be $\bar{\sigma} = \sigma / \sqrt{n}$.
% \begin{equation}
%     \bar{\sigma} = \frac{\sigma}{\sqrt{n}}
% \end{equation}

\subsection{Selecting Weights for Minimal Uncertainty}\label{solve_weight_by_noise}

When selecting weights for gyroscopes, or for accelerometers under the condition that the selection of vehicle frame is flexible, we can minimize the output noise or bias drift rate. Below, we take the bias drift as an example but similar ideas can be applied to the output noise.

The combined drift rate is as follows.
\begin{equation}
    \dot{\bar{\textbf{b}}}(t) = \sum_i{w_i \textbf{C}_i \textbf{w}_i(t)}.
\end{equation}
